---
title: "HW 4"
author: "Andrew Plantinga"
date: "2023-10-31"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(nloptr)
library(knitr)
```


## Exhaustible Resource Extraction {.tabset}


Suppose there is an exhaustible resource that is costlessly extracted.  The inverse demand curve for the resource is  $p_t=a-bq_t$ where $q_t$ is the quantity extracted in time t and a=40, b=1.  The discount rate is $\delta=0.05$  and the initial stock of reserves is $R_0$ = 778.1259 units.  Set up a program to find the competitive market equilibrium and show the equilibrium values of $p_t$, $q_t$, $R_t$, and  $\pi_t$ in a table.  $R_t$ is the reserves remaining in year t and $\pi_t$ is the discounted profits in each year (without costs, they are just equal to $\rho^tp_tq_t$ ). [Hint: it is easiest to start with the price in the last period and work backwards through time.]

### A. 

List the values of $t$, $p_t$, $q_t$, $R_t$, and  $\pi_t$ in a table below.  What should be the final (year T) values of $p_T$ and $R_T$? Why?  Does the resource get more scarce over time?  What is your evidence?

<span style="color: blue;">**Solution**</span>

```{r}
exhaust <- function( T, a, b, delta, R0){
  
  t=seq(0,T)
  
  price <- vector(mode="numeric", length = length(t))
  quantity <- vector(mode="numeric", length = length(t))
  reserves <- vector(mode="numeric", length = length(t))
  profits <- vector(mode="numeric", length = length(t))
  
  price[T+1] = a
  reserves[1] = R0
  
  for(i in (T + 1):2){
    
    quantity[i] = (a-price[i])/b 
    price[i-1] = a - b*quantity[i]
  }
    
  quantity[1] = (a-price[1])/b
  #reserves[1] = R0
  profits[1] = price[1] * quantity[1]
  #price[1] = a-b*quantity[1] #is this needed??
  #price[1] = a/(1+delta)
  
  for(i in 2:(T+1)){
    
    reserves[i] = reserves[i-1] - quantity[i-1] #do you add or subtract?
    profits[i] = 1/(1+delta) * price[i] * quantity[i]
    #price[i] = a/(1+delta) #not sure if this is right
    
  }
  
  return(data.frame("Time" = t, "Price"=price, "Quantity"=quantity, "Reserves" = reserves, "Profits" = profits)) #t, p, q, R, and pi, do you need to define T?
  
}

table1a <- exhaust(T= 10, a= 40, b= 1, delta = 0.05, R0= 778.1259)
table1a
```




### B.

Now consider extensions of the model in A.  First, assume there is a constant marginal cost of extraction c = 3.178025.  The discount profits in each year are now $\pi_t=\rho^t (p_t-c)q_t$.  Second, assume marginal costs are zero again but the initial stock of reserves increases to $R_0$ = 948.2253 units.  Third, assume marginal costs are still zero and reserves go back to $R_0$=778.1259, but the demand curve parameters are now $a$=40, $b$=0.75194.  A lower value of $b$ means that the demand curve rotates outward around $a$=40. 

List the values of $t$, $p_t$, $q_t$, $R_t$, and  $\pi_t$ for the three cases. 


<span style="color: blue;">**Solution**</span>

<span style="color: blue;">

</span>

```{r}

```


### C.


What is the final year $T$ for the three cases? Explain why they are different from part A.

<span style="color: blue;">**Solution**</span>


<span style="color: blue;">

</span>

### D. 

Compare the path of prices in the three cases to that in part A. Explain the differences.

<span style="color: blue;">**Solution**</span>

<span style="color: blue;">

</span>


### E.

What happens to the present discounted value of profits (in other words, the value of the reserves $R_0$) in the three cases compared to part A? Explain the differences.

<span style="color: blue;">**Solution**</span>

<span style="color: blue;">

</span>



