---
title: "HW 4"
author: "Kat Aristi"
date: "2023-11-8"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(nloptr)
library(knitr)
```


## Exhaustible Resource Extraction {.tabset}


Suppose there is an exhaustible resource that is costlessly extracted.  The inverse demand curve for the resource is  $p_t=a-bq_t$ where $q_t$ is the quantity extracted in time t and a=40, b=1.  The discount rate is $\delta=0.05$  and the initial stock of reserves is $R_0$ = 778.1259 units.  Set up a program to find the competitive market equilibrium and show the equilibrium values of $p_t$, $q_t$, $R_t$, and  $\pi_t$ in a table.  $R_t$ is the reserves remaining in year t and $\pi_t$ is the discounted profits in each year (without costs, they are just equal to $\rho^tp_tq_t$ ). [Hint: it is easiest to start with the price in the last period and work backwards through time.]

### A. 

List the values of $t$, $p_t$, $q_t$, $R_t$, and  $\pi_t$ in a table below.  What should be the final (year T) values of $p_T$ and $R_T$? Why?  Does the resource get more scarce over time?  What is your evidence?

<span style="color: blue;">**In the final year T=36, $p_t$=40, and $R_t$=0. Since these reserves contain an exhaustible resource, they will eventually end up being depleted and reach 0, which in this case occurs at year 36 due to the current set parameters. Therefore, the reserves and quantity slowly decrease as time increases. When reserves and quantity reach 0, profits will also be 0 since there is nothing left to make money off of. **</span>
**</span>

```{r}
exhaust <- function( T, a, b, c, delta, R0){
  
  t=seq(0,T)
  
  price <- vector(mode="numeric", length = length(t))
  quantity <- vector(mode="numeric", length = length(t))
  reserves <- vector(mode="numeric", length = length(t))
  profits <- vector(mode="numeric", length = length(t))
  
  price[T+1] = a
  reserves[1] = R0
  
  for(i in (T + 1):2){
    
    quantity[i] = (a-price[i])/b 
    price[i-1] = (price[i] + c*delta)/(1+delta)
  
  }
    
  quantity[1] = (a-price[1])/b
  profits[1] = price[1] * quantity[1]
  
  for(i in 2:(T+1)){
    
    reserves[i] = reserves[i-1] - quantity[i-1] 
    profits[i] = 1/(1+delta) * price[i] * quantity[i]
    
  }
  
  return(data.frame("Time" = t, "Price"=price, "Quantity"=quantity, "Reserves" = reserves, "Profits" = profits)) 
  
}

table1a <- exhaust(T= 36, a= 40, b= 1, delta = 0.05, R0= 778.1259, c=0)
kable(table1a)
```




### B.

Now consider extensions of the model in A.  First, assume there is a constant marginal cost of extraction c = 3.178025.  The discount profits in each year are now $\pi_t=\rho^t (p_t-c)q_t$.  Second, assume marginal costs are zero again but the initial stock of reserves increases to $R_0$ = 948.2253 units.  Third, assume marginal costs are still zero and reserves go back to $R_0$=778.1259, but the demand curve parameters are now $a$=40, $b$=0.75194.  A lower value of $b$ means that the demand curve rotates outward around $a$=40. 

List the values of $t$, $p_t$, $q_t$, $R_t$, and  $\pi_t$ for the three cases. 


<span style="color: blue;">**See tables below!**</span>

<span style="color: blue;">

</span>

Case 1:
```{r}
exhaust_case1 <- function( T, a, b, c, delta, R0){
  
  t=seq(0,T)
  
  price <- vector(mode="numeric", length = length(t))
  quantity <- vector(mode="numeric", length = length(t))
  reserves <- vector(mode="numeric", length = length(t))
  profits <- vector(mode="numeric", length = length(t))
  
  price[T+1] = a
  reserves[1] = R0
  
  for(i in (T + 1):2){
    
    quantity[i] = (a-price[i])/b 
    price[i-1] = (price[i] + c*delta)/(1+delta)
  }
    
  quantity[1] = (a-price[1])/b
  profits[1] =  delta^t * (price[1] - c) * quantity[1]
  
  for(i in 2:(T+1)){
    
    reserves[i] = reserves[i-1] - quantity[i-1] 
    profits[i] = delta^t * price[i] * quantity[i]
    
  }
  
  return(data.frame("Time" = t, "Price"=price, "Quantity"=quantity, "Reserves" = reserves, "Profits" = profits)) 
  
}

table1b_case1 <- exhaust_case1(T= 38, a= 40, b= 1, delta = 0.05, R0= 778.1259, c=3.178025)
kable(table1b_case1)
```
Case 2: 
```{r}
table1b_case2 <- exhaust_case1(T= 41, a= 40, b= 1, delta = 0.05, R0= 948.2253, c=0)
kable(table1b_case2)
```
Case 3: 
```{r}
table1b_case3 <- exhaust_case1(T= 30, a= 40, b= 0.75194, delta = 0.05, R0= 778.1259, c=0)
kable(table1b_case3)
```


### C.


What is the final year $T$ for the three cases? Explain why they are different from part A.

<span style="color: blue;">**The final T for Case One is T=38, Case Two T=41, and Case Three T=30. They are different from part A due to the change in parameters. Beginning with the change of cost in Case 1 from 0 to 3.178025. This increase in cost caused time to increase slightly from 36 to 38. This increase in cost caused both profits and prices to increase and reserves to reach 0 at a slower rate than seen in part A. When costs are higher there is an incentive to push costs off to the future and increase the time it takes for reserves to reach 0, therefore having a longer T in order to reach the price which was used to maintain the initial reserve size. In Case 2, T increases to 41 due to an increase in the amount of reserves, therefore it takes longer to reach 0 reserves. In Case 3, T decreases to 30 due to b decreasing to 0.75194. When the demand is decreased, initial price increases, so in order to ensure that you maintain the same amount of reserves your T curve (the northeast quadrant) has to become steeper. **</span>


<span style="color: blue;">

</span>

```{r}
cases <- 1:3
T <- c(38, 41, 30)
one_c <- data.frame('Cases' = cases, 'Time' = T)
kable(one_c)
```

### D. 

Compare the path of prices in the three cases to that in part A. Explain the differences.

<span style="color: blue;">**The price paths of the three cases are either decreasing at a faster or slower rate than the price path in part A. The price path of part A is between that of case 1 and 2 due to the amount of time it takes for the price to reach 0, or when there are 0 reserves (as seen in the graphs below). Case 3 has the longest price path due to having the longest amount of time for reserves to reach 0. The changes in the price paths between the three cases and that in part A can possibly be changed by either addign costs or increasing reserves. This will change the amount of time it takes these prices to go from 40 to 0.   **</span>

<span style="color: blue;">

</span>

```{r}
plot(table1b_case1$Quantity, table1b_case1$Price)
plot(table1b_case2$Quantity, table1b_case2$Price)
plot(table1b_case3$Quantity, table1b_case3$Price)

plot(table1a$Quantity, table1a$Price)
```


### E.

What happens to the present discounted value of profits (in other words, the value of the reserves $R_0$) in the three cases compared to part A? Explain the differences.

<span style="color: blue;">**Across all cases, the profits are increasing. The cases have some profit when initially starting in time 0 due to the fact that there are already exhaustible resources in the reserve that are ready to be extracted and sold. PVNB for Case 1 and 2 are the slightly larger than part A due to the time spans being longer, which allows for a longer amount of time for the profits to increase. Case 3 is the largest because the initial price is the largest causing profits to increase. As the profits decrease, they decrease at a much sharper rate than they increase due to the because the exhaustible resource cannot renew itself so is growing smaller at a fast rate. This is further seen as an increase in prices in an attempt to increase profits, but the reserves continue to decline.  **</span>

<span style="color: blue;">

</span>

```{r}
c1 <- sum(table1b_case1$Profits)
c2 <- sum(table1b_case2$Profits)
c3 <- sum(table1b_case3$Profits)

question1a <- sum(table1a$Profits)

number <- c("1A", 'Case 1', 'Case 2', 'Case 3')
PVNB <- c(question1a, c1, c2, c3)

question1e <- data.frame("Case" = number, 'PVNB' = PVNB)

kable(question1e)
```


