---
title: "ESM 242 Midterm Exam"
author: "Kat Aristi"
output: 
  html_document:
    code_folding: hide
date: '2023-11-08'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(nloptr)
library(knitr)
```








## Question 1: Stand Externalities {.tabset}

This problem considers spatial interactions between two stands.  Suppose that the management of stand 1 affects the growth of stand 2.  This is a classic case of an externality. Depending on the specific setting, this could be a negative or positive externality whereby stand 1 reduces or increases the growth of stand 2.  I ask you to consider how the externality affects the solution to the optimal rotation length for stand 2 and to think about how policies could be designed to remedy the problem.

Suppose the yield functions for stand 1 and 2 are, respectively, $Q(T_1)=e^{a-b/T_1}$  and $Q(T_2,T_1^* )=e^{(a-b/T_2+dT_1^*} )$. The yield for stand 1 is the usual specification, but for stand 2, the yield depends on the age of the stand as well as the rotation length chosen for the first stand ($T_1^*$ ).  Note that the yield function for stand 2 depends on the rotation length $T_1^*$  and not on the current age of the stand $T_1$ .  While the latter formulation is more realistic, it is much harder to implement in this problem.  Define a=10, b=90, c=125, p=2.5, and $\delta$=0.03.

### A.

Using NLoptR and the parameter values given above, what is the optimal rotation length for stand 1 ( $T_1^*$) when the stand is grown for an infinite number of rotations?

<span style="color: blue;">**The optimal rotation length for stand 1 is 48.25787.**</span>



```{r}
obj_fcn_infinite_1a <- function(choice,a,b,c,p,discount){

  Q=exp(a-b/choice)
  benefits = (p*Q*exp(-discount*choice)-c)/(1-exp(-discount*choice))
  
  return(-benefits)
  
}

options=list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15,maxeval=16000)

out_tree1a=nloptr(x0=10,
          eval_f=obj_fcn_infinite_1a,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=90,c=125,p=2.5)

#print ( out_tree1a$solution )
#print(-out_tree1a$objective)

```




### B.

Holding the rotation for stand 1 at the value of $T_1^*$  found in a, what are the optimal rotations (assume an infinite number of them) for stand 2 ($T_2^*$) for three different values of d: d=0, d=-0.03, d=0.1?  Explain the relationship between $T_2^*$ and $d$.

<span style="color: blue;">**Solution**.</span>



```{r}
obj_fcn_infinite_1b <- function(choice,a,b,c,p,d,discount){
  
  Q=exp(a-b/choice + d*48.25787)
  benefits = (p*Q*exp(-discount*choice)-c)/(1-exp(-discount*choice))
  
  return(-benefits)
}

options=list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15,maxeval=16000)

out_tree1b=nloptr(x0=10,
          eval_f=obj_fcn_infinite_1b,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=90,c=125,p=2.5, d=0)

print(out_tree1b$solution)
```

```{r}
out_tree1bb=nloptr(x0=10,
          eval_f=obj_fcn_infinite_1b,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=90,c=125,p=2.5, d=-0.03)

print(out_tree1bb$solution)
```

```{r}
out_tree1bc=nloptr(x0=10,
          eval_f=obj_fcn_infinite_1b,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=90,c=125,p=2.5, d=0.1)

print(out_tree1bc$solution)
```
```{r}
d <- c(0, -0.03, 0.1)
rotation <- c(48.25787, 49.70415, 47.80502)

question1b <- data_frame("D" = d, "Rotation Length" = rotation)

kable(question1b)
```


### C.

What are some real-world examples of negative and positive externalities that affect the growth of neighboring stands?  Describe some policy interventions that could be used to get the owner of stand 1 to internalize the effect of her stand on stand 2.  In order words, how could policies be used to get the owner of stand 1 to reduce negative externalities or increase positive externalities?  

<span style="color: blue;">**Solution**.</span>





## Question 2: Stock Pollutant {.tabset}

This question builds on the stock pollutant example presented in class. The objective is to maximize the present value of net benefits (PVNB) from production $q_t$ of a good: 

$$
PVNB= \max_q \sum^{t=10}_{t=0}\rho^t(pq_t-(aZ_t-bZ_t^2))
$$
where $(aZ_t-bZ_t^2)$ are the damages from the stock pollution $Z_t$.  The maximization problem is subject to $Z_{t+1}=Z_t(1-\gamma)+\alpha*q_t$ and $Z_0$ given. Use the following parameter values: $\rho=\frac{1}{1.04}$, p=100, a= 2, b=-0.25,$\gamma=0.2$, $\alpha=2$, and $Z_0$=40. In addition, I want you to ensure that the pollution stock in the last period ($Z_{10}$) is equal to 25. You can think of $Z_{10}$ as a stabilization target, which mimics the goals of the Paris Agreement ( https://unfccc.int/process-and-meetings/the-paris-agreement/the-paris-agreement).  To implement $Z_{10}=25$, you will need a constraint just like the one we used to ensure that $q_t \leq R_t$ (HW 1) and $Y_t \leq X_t$ (HW 2).  However, in this case you only need to constrain the final ($t=10$) value.  An equality constraint requires different options than the inequality constraint.  I provide these for you below.  In your nloptr command, you will need to call the constraint with eval_g_eq=name_of_constraint. I also suggest including the lower bound constraint lb=rep(0,period) to ensure that extract is non-negative.    


### A. 

What is the optimal solution? Make a table reporting $t$, $q_t$, $Z_t$, and $\pi_t=\rho^t(pq_t-(aZ_t-bZ_t^2))$.

<span style="color: blue;">**Solution**
</span>



```{r}




  
  
 



#local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
#options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)

#out_3a=nloptr(
#              eval_g_eq= ,
#              opts = options,
              
#              )




```









### B. 

How does your answer to part a change when the price parameter p is raised to 150?  Explain why the optimal $q_t$ changes compared to part a.

<span style="color: blue;">**Solution**
</span>




### C.

Set the price back to 100. How does your answer to part a change when the stabilization target is lowered to $Z_{11} =10$?  Explain the change in $q_t$.  

<span style="color: blue;">**Solution**
</span>



### D.

Why do you think the pollution stock in period 5, $Z_5$, is virtually the same under the solutions in parts a and c?

<span style="color: blue;">**Solution**
</span>

