---
title: "ESM 242 Midterm Exam"
author: "Kat Aristi"
output: 
  html_document:
    code_folding: hide
date: '2023-11-15'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(nloptr)
library(knitr)
```








## Question 1: Stand Externalities {.tabset}

This problem considers spatial interactions between two stands.  Suppose that the management of stand 1 affects the growth of stand 2.  This is a classic case of an externality. Depending on the specific setting, this could be a negative or positive externality whereby stand 1 reduces or increases the growth of stand 2.  I ask you to consider how the externality affects the solution to the optimal rotation length for stand 2 and to think about how policies could be designed to remedy the problem.

Suppose the yield functions for stand 1 and 2 are, respectively, $Q(T_1)=e^{a-b/T_1}$  and $Q(T_2,T_1^* )=e^{(a-b/T_2+dT_1^*} )$. The yield for stand 1 is the usual specification, but for stand 2, the yield depends on the age of the stand as well as the rotation length chosen for the first stand ($T_1^*$ ).  Note that the yield function for stand 2 depends on the rotation length $T_1^*$  and not on the current age of the stand $T_1$ .  While the latter formulation is more realistic, it is much harder to implement in this problem.  Define a=10, b=90, c=125, p=2.5, and $\delta$=0.03.

### A.

Using NLoptR and the parameter values given above, what is the optimal rotation length for stand 1 ( $T_1^*$) when the stand is grown for an infinite number of rotations?

<span style="color: blue;">**The optimal rotation length for stand 1 is 48.25787.**</span>



```{r}
obj_fcn_infinite_1a <- function(choice,a,b,c,p,discount){

  Q=exp(a-b/choice)
  benefits = (p*Q*exp(-discount*choice)-c)/(1-exp(-discount*choice))
  
  return(-benefits)
  
}

options=list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15,maxeval=16000)

out_tree1a=nloptr(x0=10,
          eval_f=obj_fcn_infinite_1a,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=90,c=125,p=2.5)

#print ( out_tree1a$solution )
#print(-out_tree1a$objective)

```




### B.

Holding the rotation for stand 1 at the value of $T_1^*$  found in a, what are the optimal rotations (assume an infinite number of them) for stand 2 ($T_2^*$) for three different values of d: d=0, d=-0.03, d=0.1?  Explain the relationship between $T_2^*$ and $d$.

<span style="color: blue;"> **d is an externality that is occuring in stand 1 and effecting the rotation length of stand 2. When d=0, there is no externality occurring. When d=-0.3, rotation length in stand 2 increases and when d=0.1 stand 2s rotation length decreases.**.</span>



```{r}
obj_fcn_infinite_1b <- function(choice,a,b,c,p,d,discount){
  
  Q=exp(a-(b/choice) + (d*48.25787)) 
  benefits = (p*Q*exp(-discount*choice)-c)/(1-exp(-discount*choice))
  
  return(-benefits)
}

options=list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15,maxeval=16000)

out_tree1b=nloptr(x0=10,
          eval_f=obj_fcn_infinite_1b,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=90,c=125,p=2.5, d=0)

#print(out_tree1b$solution)
```

```{r}
out_tree1bb=nloptr(x0=10,
          eval_f=obj_fcn_infinite_1b,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=90,c=125,p=2.5, d=-0.03)

#print(out_tree1bb$solution)
```

```{r}
out_tree1bc=nloptr(x0=10,
          eval_f=obj_fcn_infinite_1b,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=90,c=125,p=2.5, d=0.1)

#print(out_tree1bc$solution)
```

```{r}
d <- c(0, -0.03, 0.1)
rotation <- c(48.25787, 49.70415, 47.80502)

question1b <- data_frame("D" = d, "Rotation Length" = rotation)

kable(question1b)
```


### C.

What are some real-world examples of negative and positive externalities that affect the growth of neighboring stands?  Describe some policy interventions that could be used to get the owner of stand 1 to internalize the effect of her stand on stand 2.  In order words, how could policies be used to get the owner of stand 1 to reduce negative externalities or increase positive externalities?  

<span style="color: blue;">**Let's say that stand 1 is a privately owned tree stand managed for logging purposes while stand 2 is an adjacent privately owned tree stand managed for ecological conservation. An example of a negative externality could be that the logging practices in stand 1 generates run off that affects the growth of the trees in stand 2 as well as the habitat for the wildlife and biodiversity of the trees in stand 2. A potential policy intervention could be implementing regulations or standards for the logging practices occuring in stand 1. This could involve specific run off management, such as controlling the sediment and chemical run off. Stand 1 may have to retain certain vegetation in their stand to prevent the run off and stabilize the soil. An example of a positive externality could be stand 1 engaging in sustainable logging practices that enhance soil quality. These practices could be selective cutting or low impact logging that causes minimal disruption to the soil as well as leaving organic matter on the floor of the forest which improves the soil quality. This could then help improve the soil quality in stand 2 and contribute to the overall growth of the stand and ecological conservation in the stand due to the soil not being disturbed. A policy intervention that could increase this positive externality would be providing financial incentives or subsidies to the owner of stand 1 for adopting and maintaining sustainable logging practices. This could be a tax break for eco-friendly certifications.  **.</span>



## Question 2: Stock Pollutant {.tabset}

This question builds on the stock pollutant example presented in class. The objective is to maximize the present value of net benefits (PVNB) from production $q_t$ of a good: 

$$
PVNB= \max_q \sum^{t=10}_{t=0}\rho^t(pq_t-(aZ_t-bZ_t^2))
$$
where $(aZ_t-bZ_t^2)$ are the damages from the stock pollution $Z_t$.  The maximization problem is subject to $Z_{t+1}=Z_t(1-\gamma)+\alpha*q_t$ and $Z_0$ given. Use the following parameter values: $\rho=\frac{1}{1.04}$, p=100, a= 2, b=-0.25,$\gamma=0.2$, $\alpha=2$, and $Z_0$=40. In addition, I want you to ensure that the pollution stock in the last period ($Z_{10}$) is equal to 25. You can think of $Z_{10}$ as a stabilization target, which mimics the goals of the Paris Agreement ( https://unfccc.int/process-and-meetings/the-paris-agreement/the-paris-agreement).  To implement $Z_{10}=25$, you will need a constraint just like the one we used to ensure that $q_t \leq R_t$ (HW 1) and $Y_t \leq X_t$ (HW 2).  However, in this case you only need to constrain the final ($t=10$) value.  An equality constraint requires different options than the inequality constraint.  I provide these for you below.  In your nloptr command, you will need to call the constraint with eval_g_eq=name_of_constraint. I also suggest including the lower bound constraint lb=rep(0,period) to ensure that extract is non-negative.    


### A. 

What is the optimal solution? Make a table reporting $t$, $q_t$, $Z_t$, and $\pi_t=\rho^t(pq_t-(aZ_t-bZ_t^2))$.

<span style="color: blue;">**Solution**
</span>


```{r}
obj_fcn_stock<-function(choice,int_state,a,b,gamma,p,alpha,discount){

  state<-vector(mode="numeric",length=length(choice))
  benefits<-vector(mode="numeric",length=length(choice))
  
  state[1]=int_state
  benefits[1]=p*choice[1]-(a*state[1]-b*state[1]^2)
    
  for(i in 2:length(choice)){

  state[i]=state[i-1]*(1-gamma)+alpha*choice[i-1] 
  benefits[i]=p*choice[i]-(a*state[i]-b*state[i]^2)
       
  }
  
  t=seq(from=0,length.out=length(benefits))
  pv=discount^t*benefits
  npv=sum(pv)
  return(-npv)
  
}


constraint_function <- function(choice, int_state, a, b, gamma, p, alpha, discount){
  
  state<-vector(mode="numeric",length=period)
  
  state[1] = int_state

  for (i in 2:length(choice)) {
    state[i] = state[i - 1] * (1 - gamma) + alpha * choice[i - 1]
  }
  
  const = state[11] - 25
  return(const)
}
period = 11

local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)
out_2a=nloptr(x0=rep(1, period),
             eval_f = obj_fcn_stock,
             eval_g_eq = constraint_function ,
             opts = options,
             lb = rep(0, period),
             int_state = 40,
             discount = 1/1.04,
             a = 2,
             b = -.25,
             gamma = 0.2,
             alpha = 2,
             p = 100)

#round(out_2a$solution, 2)

```

```{r}
period = 11
stock_2a <- function(choice, int_state=40, period=11, a=2, b=-0.25, p=100, alpha=2, gamma=0.2, discount=1/1.04){
  
  state<-vector(mode="numeric",length=0)
  benefits<-vector(mode="numeric",length=0)
  
  state[1] = int_state
  benefits[1] = p*choice[1]-(a*state[1]-b*state[1]^2)
  
  for(i in 2:period){
    
    state[i]=state[i-1]*(1-gamma)+alpha*choice[i-1]
    benefits[i]=p*choice[i]-(a*state[i]-b*state[i]^2)
    
  }
  
  #state[11]=final_state
  
  t=seq(from=0, length.out=length(benefits))
  pv=discount^t*benefits
  data.frame("Period"=t, "PVNB"=pv, "Stock Pollution"=state, "Production of a Good"=choice)
  
}

#out_2a$solution[11] = NA

tbl_2a=round(stock_2a(out_2a$solution),2)    


kable(tbl_2a)
  
```

I am sorry my numbers are weird :(

### B. 

How does your answer to part a change when the price parameter p is raised to 150?  Explain why the optimal $q_t$ changes compared to part a.

<span style="color: blue;">**When the price parameter is raised to 150, the optimal $q_t$ should decrease. This is because it becomes more costly to produce this good and therefore more expensive to pollute. Due to this spike in cost, less of the good is produced and less pollution is emitted into the air. Price of production is lower in part a, so more goods are then produced. **
</span>

```{r}
period=11
out_2b=nloptr(x0=rep(1, period),
             eval_f = obj_fcn_stock,
             eval_g_eq = constraint_function ,
             opts = options,
             lb = rep(0, period),
             int_state = 40,
             discount = 1/1.04,
             a = 2,
             b = -.25,
             gamma = 0.2,
             alpha = 2,
             p = 150)

#round(out_2b$solution, 2)

stock_2b <- function(choice, int_state=40, period=11, a=2, b=-0.25, p=150, alpha=2, gamma=0.2, discount=1/1.04){
  
  state<-vector(mode="numeric",length=length(choice))
  benefits<-vector(mode="numeric",length=length(choice))
  
  state[1] = int_state
  benefits[1] = p*choice[1]-(a*state[1]-b*state[1]^2)
  
  for(i in 2:period){
    
    state[i]=state[i-1]*(1-gamma)+alpha*choice[i-1]
    benefits[i]=p*choice[i]-(a*state[i]-b*state[i]^2)
    
  }
  
  t=seq(from=0,length.out=length(benefits))
  pv=discount^t*benefits
  data.frame("Period"=t, "PVNB"=pv, "Stock Pollution"=state, "Production of a Good"=choice)
  
}

#out_2a$solution[11] = NA

tbl_2b=round(stock_2b(out_2b$solution), 2)    


kable(tbl_2b)
```



### C.

Set the price back to 100. How does your answer to part a change when the stabilization target is lowered to $Z_{10} =10$?  Explain the change in $q_t$.  

<span style="color: blue;">**When the stabilization target is lowered to $Z_{10}=10$, there is a stricter cap being put on the amount of pollution that is being allowed to be released into the atmosphere. Since producers now need to pollute less than they were originally doing in a, less of the good is able to be produced causing $q_t$ to decrease. **
</span>


```{r}
period=11
obj_stock_func_2c <- function(choice, int_state, p, a, b, gamma, alpha, discount){ 
  
  state<-vector(mode="numeric", length=length(choice))
  benefits<-vector(mode="numeric", length=length(choice))
  
  state[1] = int_state
  benefits[1] = p*choice[1]-(a*state[1]-b*state[1]^2) 

  
  for(i in 2:length(choice)){
    
    state[i]=state[i-1]*(1-gamma)+alpha*choice[i-1]
    benefits[i]=p*choice[i]-(a*state[i]-b*state[i]^2)
    
  }
  
  t=seq(from=0,length.out=length(benefits))
  pv=discount^t*benefits
  npv=sum(pv)
  return(-npv)
  
}

constraint_function_2c <- function(choice, p, a, b, gamma, alpha, discount, int_state){
  
  state <-vector(mode="numeric",length=period)
  #const <-vector(mode="numeric",length=0)
  
  state[1] = int_state 
  #const <- state[11]-25
  
  for(i in 2:length(choice)){
    
    state[i] = state[i-1]*(1-gamma) + alpha*choice[i-1]
    #const[i] = state[11]-25 
    
  }
  const=state[11]-10
  
  return(const)
}


local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)

period = 11

out_2c=nloptr(x0 = rep(1, period),
              int_state = 40, 
              eval_f = obj_stock_func_2c,
              eval_g_eq= constraint_function_2c,
              lb = rep(0, period),
              opts = options,
              discount = 1/1.04, 
              p=100,
              a=2,
              b=-0.25,
              gamma=0.2, 
              alpha=2
              )

stock_2c <- function(choice, int_state=40, period=10, a=2, b=-0.25, p=150, alpha=2, gamma=0.2, discount=1/1.04){
  
  state<-vector(mode="numeric",length=length(choice))
  benefits<-vector(mode="numeric",length=length(choice))
  
  state[1] = int_state
  benefits[1] = p*choice[1]-(a*state[1]-b*state[1]^2)
  
  for(i in 2:length(choice)){
    
    state[i]=state[i-1]*(1-gamma)+alpha*choice[i-1]
    benefits[i]=p*choice[i]-(a*state[i]-b*state[i]^2)
    
  }
  
  t=seq(from=0,length.out=length(benefits))
  pv=discount^t*benefits
  data.frame("Period"=t, "PVNB"=pv, "Stock Pollution"=state, "Production of a Good"=choice)
  
}

#out_2a$solution[11] = NA

tbl_2c=round(stock_2a(out_2b$solution),2)    


kable(tbl_2c)
```


### D.

Why do you think the pollution stock in period 5, $Z_5$, is virtually the same under the solutions in parts a and c?

<span style="color: blue;">**$Z_5$ in parts a and c should be the same because only the final period, T10, is being constrained. Since a and c are being constrained at different amounts in their final period, only that one should differ causing the other periods pollution stock to remain the same between the two questions. **
</span>
